A Hybrid Ensemble Learning Approach to Star-galaxy Classification

### 2015-04-13 paper/main.tex
add purity vs mag plot.

\begin{figure*}
  \begin{minipage}[b]{0.49\linewidth}
    \includegraphics[width=\textwidth]{figures/som_colors.eps}
    \caption{A two-dimensional 10$\times$10 SOM representation
             showing the mean $i$-band magnitude (top left),
             the fraction of true stars in each cell (top right),
             and the mean values of $u-g$ and $g-r$
             for the cross validation data.}
    \label{fig:som_colors}
  \end{minipage}
  \hfill
  \begin{minipage}[b]{0.49\linewidth}
    \includegraphics[width=\textwidth]{figures/weights.eps}
    \caption{A two-dimensional 10$\times$10 SOM representation
             showing the relative weights for the BMC combination technique
             applied to the four individual methods for the CFHTLenS data.}
    \label{fig:weights}
  \end{minipage}
\end{figure*}

@@

We can also use the same SOM to determine the relative weights for 
the four individual classification methods for each cell.
We present the four weight maps for the BMC technique
in Figure~\ref{fig:weights}.
In these maps, a redder color indicates a higher weight,
or equivalently that the corresponding weight performs better in that region.
These weight maps demonstrate the variation in
the performance of the individual techniques across
the two-dimensional parameter space defined by the SOM.
Not surprisingly, the morphological separation
performs best in the top left corner of the $i$-band magnitude map,
which corresponds to the brightest CFHTLenS magnitudes $i \lesssim 20$.
On the other hand, \texttt{TPC} seems to perform best
in the region that corresponds to intermediate magnitudes
$20\gtrsim i \lesssim22.5$ and $1.5<u-g<3.0$.
Our unsupervised learning method \texttt{SOMc}
performs relatively better at fainter magnitudes $i>21.5$
with $0<u-g<0.5$ and $0<g-r<0.5$.
\textcolor{blue}{
Although \texttt{HB} shows the worst performance
when there exists ample training data,
BMC still utilizes information from \texttt{HB}
to produce an overall improvement in performance,
especially at intermediate magnitudes $20\gtrsim i \lesssim20$.
}

@@

It is very costly in terms of telescope time to
obtain a large sample spectroscopic observations
down to the limiting magnitude of a photometric sample.
Thus, we investigate the impact of training set quality
by gradually decreasing the size of the training set.

\begin{figure*}
  \begin{minipage}[b]{0.49\linewidth}
    \includegraphics[width=\textwidth]{figures/perform_mag_cut.eps}
    \caption{The classification performance metrics for TPC, HB, morphology,
      and BMC as applied to the CFHTLenS data in the VVDS field
      with various magnitude cuts.
      We show only one of the four combination methods, BMC,
      which has the best overall performance.}
    \label{fig:perform_mag_cut}
  \end{minipage}
  \hfill
  \begin{minipage}[b]{0.49\linewidth}
    \includegraphics[width=\textwidth]{figures/perform_z_cut.eps}
    \caption{Same as Figure.~\ref{fig:perform_mag_cut} using
      $z_{\text{spec}}$ cuts.}
    \label{fig:perform_z_cut}
  \end{minipage}
\end{figure*}

@@

We also investigate the impact of training set quality
by analyzing in detail a more realistic case
where the training data is available
only for a small number of objects at bright magnitudes.
To emulate this scenario,
we only use objects that have spectroscopic labels
from the VVDS 0226-04 field (which is located within the CFHTLS W1 field)
and impose a magnitude cut of $i < 22.0$ in the training data,
leaving us a training set with 6,361 objects.

@@

\begin{figure*}
  \begin{minipage}[t]{0.49\linewidth}
    \includegraphics[width=\textwidth]{figures/purity_mag.eps}
    \caption{Purity of morphological separation,
             TPC, and BMC techniques in 
             Table~\ref{table:metrics_vvds}
             as a function of the $i$-band magnitude.
             The top panel shows the fraction of stars
             as a function of magnitude.
             the galaxy and star purity for TPC and BMC techniques.
             The bottom two panels show
             the galaxy and star purity for morphological separation
             and BMC technique.
             We use the same threshold values from 
             Figure~\ref{fig:purity_mag} in the bottom four panels}.
    \label{fig:purity_mag}
  \end{minipage}
  \hfill
  \begin{minipage}[t]{0.49\linewidth}
    \includegraphics[width=\textwidth]{figures/purity_z.eps}
    \caption{The same figures as Figure~\ref{fig:purity_mag},
             but as a function of photometric redshift estimate
             (photo-$z$).}
    \label{fig:purity_z}
  \end{minipage}
\end{figure*}

>>>>

\begin{figure*}
  \begin{minipage}[t]{0.49\linewidth}
    \includegraphics[width=\textwidth]{figures/som_colors.eps}
    \caption{A two-dimensional 10$\times$10 SOM representation
             showing the mean $i$-band magnitude (top left),
             the fraction of true stars in each cell (top right),
             and the mean values of $u-g$ and $g-r$
             for the cross validation data.}
    \label{fig:som_colors}
  \end{minipage}
  \hfill
  \begin{minipage}[t]{0.49\linewidth}
    \includegraphics[width=\textwidth]{figures/weights.eps}
    \caption{A two-dimensional 10$\times$10 SOM representation
             showing the relative weights for the BMC combination technique
             applied to the four individual methods for the CFHTLenS data.}
    \label{fig:weights}
  \end{minipage}
\end{figure*}

@@

We can also use the same SOM to determine the relative weights for 
the four individual classification methods for each cell.
We present the four weight maps for the BMC technique
in Figure~\ref{fig:weights}.
In these maps, a redder color indicates a higher weight,
or equivalently that the corresponding weight performs better in that region.
These weight maps demonstrate the variation in
the performance of the individual techniques across
the two-dimensional parameter space defined by the SOM.
Not surprisingly, the morphological separation
performs best in the top left corner of the $i$-band magnitude map,
which corresponds to the brightest CFHTLenS magnitudes $i \lesssim 20$.
On the other hand, \texttt{TPC} seems to perform best
in the region that corresponds to intermediate magnitudes
$20\gtrsim i \lesssim22.5$ and $1.5<u-g<3.0$.
Our unsupervised learning method \texttt{SOMc}
performs relatively better at fainter magnitudes $i>21.5$
with $0<u-g<0.5$ and $0<g-r<0.5$.
\textcolor{blue}{
Although \texttt{HB} shows the worst performance
when there exists ample training data,
BMC still utilizes information from \texttt{HB}
to produce an overall improvement in performance,
especially at intermediate magnitudes $20\gtrsim i \lesssim22$.
}

In Figure~\ref{fig:purity_mag}, 
we compare the star and galaxy purity values
for morphological separation, TPC, and BMC techniques
as a function of $i$-band magnitude.
Although morphological separation shows a slightly better performance
at bright magnitudes $i < 21$,
BMC outperforms both morphological separation and TPC
at faint magnitudes $i > 21$.
As shown in the histogram, the histogram peaks at $i \sim 22$,
and BMC therefore outperforms the morphological separation technique
for the majority of objects.
It is also clear that BMC outperforms TPC over all magnitudes.
BMC can accomplish this by combining information from
all base classifiers,
\eg giving more weight to the morphological separation method
at bright magnitudes.

\begin{figure}
%%  \begin{minipage}[t]{0.49\linewidth}
    \includegraphics[width=\textwidth]{figures/purity_mag.eps}
    \caption{Purity of morphological separation,
             TPC, and BMC techniques in 
             Table~\ref{table:metrics_vvds}
             as a function of the $i$-band magnitude.
             The top panel shows the histogram of objects in the test data set.
             The second panel shows the fraction of stars
             as a function of magnitude. The error bars are Poisson.
             The bottom two panels compares
             the galaxy and star purity for morphological separation, TPC,
             and BMC techniques as a function of magnitude.
             The error bars for completeness and purity are not shown for
             clarity, but they are approximately 10 percent.
             }.
    \label{fig:purity_mag}
\end{figure}

@@

It is very costly in terms of telescope time to
obtain a large sample spectroscopic observations
down to the limiting magnitude of a photometric sample.
Thus, we investigate the impact of training set quality
by imposing various magnitude and $z$ cuts
to decrease the size of the training set.

In Figure.~\ref{fig:perform_mag_cut}, we present
a visual comparison between different classification techniques
when various magnitude cuts are applied.

In Figure.~\ref{fig:perform_z_cut}, we also present
the result of imposing various $z$ cuts.

\begin{figure*}
  \begin{minipage}[t]{0.49\linewidth}
    \includegraphics[width=\textwidth]{figures/perform_mag_cut.eps}
    \caption{The classification performance metrics for TPC, HB, morphology,
      and BMC as applied to the CFHTLenS data in the VVDS field
      with various magnitude cuts.
      The top panel shows the number of sources in the training set
      at corresponding magnitude cuts.
      We show only one of the four combination methods, BMC,
      which has the best overall performance.}
    \label{fig:perform_mag_cut}
  \end{minipage}
  \hfill
  \begin{minipage}[t]{0.49\linewidth}
    \includegraphics[width=\textwidth]{figures/perform_z_cut.eps}
    \caption{Same as Figure.~\ref{fig:perform_mag_cut} using
      $z_{\text{spec}}$ cuts.}
    \label{fig:perform_z_cut}
  \end{minipage}
\end{figure*}

@@

We also investigate the impact of training set quality
by analyzing in detail a more realistic case
where the training data is available
only for a small number of objects at bright magnitudes.
To emulate this scenario,
we only use objects that have spectroscopic labels
from the VVDS 0226-04 field (which is located within the CFHTLS W1 field)
and impose a magnitude cut of $i < 22.0$ in the training data,
leaving us a training set with 1,365 objects.

@@

\begin{figure*}
  \begin{minipage}[t]{0.49\linewidth}
    \includegraphics[width=\textwidth]{figures/purity_mag_cut.eps}
    \caption{Purity of morphological separation,
             TPC, and BMC techniques in 
             Table~\ref{table:metrics_vvds}
             as a function of the $i$-band magnitude.
             The top panel shows the fraction of stars
             as a function of magnitude.
             the galaxy and star purity for TPC and BMC techniques.
             The bottom two panels show
             the galaxy and star purity for morphological separation
             and BMC technique.
             We use the same threshold values from 
             Figure~\ref{fig:purity_mag} in the bottom four panels}.
    \label{fig:purity_mag_cut}
  \end{minipage}
  \hfill
  \begin{minipage}[t]{0.49\linewidth}
    \includegraphics[width=\textwidth]{figures/purity_z.eps}
    \caption{The same figures as Figure~\ref{fig:purity_mag},
             but as a function of photometric redshift estimate
             (photo-$z$).}
    \label{fig:purity_z}
  \end{minipage}
\end{figure*}
